---
title: "MLee Project 2 on Google_Food_Search dataset"
output: html_document
---

## Package and Library Setup
```{r}
 install.packages("dplyr", repos = "http://cran.us.r-project.org")
 install.packages("tidyr")
 install.packages("ggplot2")
 
 library(dplyr)
 library(tidyr)
 library(ggplot2)

```

## Read Google_Food_Search dataset
```{r}
 dataset <- read.csv("seasonality-values-d2b9a3.csv")
 str(dataset)
```

## Parse 'week_id' column into 'year' & 'week' columns
```{r}

 dataset <- dataset %>% separate(week_id, c("year", "week"))
 dataset$year <- as.integer(dataset$year)
 dataset$week <- as.integer(dataset$week)
 str(dataset)

```

## Plot the trend for total search for 2004-2016
```{r}

 plotData1 <- dataset            %>%
 group_by(year)                  %>%
 summarize(count=sum(value))

 g1 <- ggplot(plotData1,aes(year,count))                        +
       geom_point(color="grey",size=4,alpha=1/2)                +
       geom_smooth(se=FALSE,method='loess')                     +
       labs(x="Year",y="Count")                                 +
       labs(title="Total Google Food Search from 2004 to 2016") +
       labs(subtitle="   Search Volume Doubles as Internet Accessiblility Increases")
 g1            

```

## Filter 2 food items from dataset
##    Avocado   - all year round super food
##    Ice Cream - seasonal food
## Add 3 new columns(mon,qtr,monCount) to foodData
```{r}
 
 foodData <- dataset %>%
             filter(id %in% c('avocado','ice-cream') )
 
 foodData <- mutate(foodData, qtr = as.factor (
             quarters( as.Date(
             paste("1",foodData$week,foodData$year,sep="-"),
             format="%w-%W-%Y") ) ) )
 
 foodData <- mutate(foodData, mon = match( months( as.Date(
             paste("1",foodData$week,foodData$year,sep="-"),
             format="%w-%W-%Y") ) ,month.name) )
   
 foodData <- mutate(foodData,monCount=NA)
 str(foodData)
  
```

## Compute monthly count 'monCount' from weekly count 'value' 
```{r}

 maxRow <- dim(foodData)[1]
 i <-1
 while (i <= maxRow) {
   foodData[i,8] <- foodData[i,5]   + foodData[i+1,5] +
                    foodData[i+2,5] + foodData[i+3,5]
   if ( i+4 > maxRow) break()
   if ( foodData[i,7] == foodData[i+4,7] ) {
          foodData[i,8] <- foodData[i  ,8] +
                           foodData[i+4,5]
          i <- i+1
   }
   i <- i+4
 } # end While loop
 
```

## Delete 'value' column & keep rows with valid 'monCount'
```{r}

 sum(is.na(foodData))
 foodData <- select(foodData, -value)
 foodData <- na.omit(foodData)
 maxRow   <- dim(foodData)[1]
 
```

## Extract super food (avocado) from foodData
## Plot yearly trend
```{r}

 superFood <- foodData %>% filter(id %in% c('avocado'))

 plotData2 <- superFood           %>%
 group_by(year)                   %>%
 summarize(count=sum(monCount))
  
 g2 <- ggplot(plotData2,aes(year,count))                    +
       geom_point(color="green",size=4,alpha=1/2)           +
       geom_smooth(size=2,se=FALSE,method="loess")          +
       labs(x="Year",y="Count")                             +
       labs(subtitle="  Search Volume Increased 5 Folds ")  +
 labs(title="Google Search on Super Food Avocado  (2004-2016)")
 g2
 
```

## Plot avocado search quarterly trend
```{r}

 g3 <- ggplot(superFood,aes(qtr,monCount))                    +
       geom_boxplot(fill="green",size=1)                      +
       labs(x="Quarters",y="Count")                           +
 labs(title="Google Search on Super Food Avocado (2004-2016)")+
 labs(subtitle="  Year Round Item - limited volume changes from seasons")
 g3
 
```

## Extract seasonal food (ice cream) from foodData
## Plot yearly trend
```{r}

 seasonalFood <- foodData %>% filter(id %in% c('ice-cream'))

 plotData4 <- seasonalFood         %>%
 group_by(year)                    %>%
 summarize(count=sum(monCount))
  
 g4 <- ggplot(plotData2,aes(year,count))                    +
       geom_point(color="purple",size=4,alpha=1/2)          +
       geom_smooth(size=2,se=FALSE,method="loess")          +
       labs(x="Year",y="Count")                             +
       labs(title="Google Search on Ice Cream (2004-2016)") +
       labs(subtitle="   Search Volume Increased 3 Folds ")
 g4
 
```

## Plot ice cream quarterly trend
```{r}

 g5 <- ggplot(seasonalFood,aes(qtr,monCount))               +
       geom_boxplot(fill="purple",size=1)                   +
       coord_cartesian(ylim=c(0,125))                       +
       labs(x="Quarters",y="Count")                         +
       labs(title="Google  Search on Ice Cream (2004-2016)")+
 labs(subtitle="   Seasonal Item - search doubles in warmer seasons")
 g5
 
```