---
title: "MLee Project 3 on t-Test on sample"
output: html_document
---

## Set up population myData dataframe DF
##        by reusing code from Project #2 (line 9-76 below)

## Package and Library Setup
```{r}
 install.packages("dplyr", repos = "http://cran.us.r-project.org")
 install.packages("tidyr")
 install.packages("ggplot2")
 
 library(dplyr)
 library(tidyr)
 library(ggplot2)
```

## Read Google_Food_Search dataset
```{r}
 dataset <- read.csv("seasonality-values-d2b9a3.csv")
 str (dataset)
```

## Parse 'week_id' column into 'year' & 'week' columns
```{r}

 dataset <- dataset %>% separate(week_id, c("year", "week"))
 dataset$year <- as.integer(dataset$year)
 dataset$week <- as.integer(dataset$week)
 str(dataset)

```

## Filter 2 food items from dataset
##    Avocado   - all year round super food
##    Ice Cream - seasonal food
## Add 3 new columns(qtr,mon,monCount) to foodData
```{r}
 
 foodData <- dataset %>%
             filter(id %in% c('avocado','ice-cream') )
 
 foodData <- mutate(foodData, qtr = as.factor (
             quarters( as.Date(
             paste("1",foodData$week,foodData$year,sep="-"),
             format="%w-%W-%Y") ) ) )
 
 foodData <- mutate(foodData, mon = match( months( as.Date(
             paste("1",foodData$week,foodData$year,sep="-"),
             format="%w-%W-%Y") ) ,month.name) )
   
 foodData <- mutate(foodData,monCount=NA)
 str(foodData)
  
```

## Compute monthly count 'monCount' from weekly count 'value' 
```{r}

 maxRow <- dim(foodData)[1]
 i <-1
 while (i <= maxRow) {
   foodData[i,8] <- foodData[i,5]   + foodData[i+1,5] +
                    foodData[i+2,5] + foodData[i+3,5]
   if ( i+4 > maxRow) break()
   if ( foodData[i,7] == foodData[i+4,7] ) {
          foodData[i,8] <- foodData[i  ,8] +
                           foodData[i+4,5]
          i <- i+1
   }
   i <- i+4
 } # end While loop
 
```

## Project #3 code starts here
## Load myData dataframe for data analysis
```{r}

 myData <- select(foodData, id, monCount)  %>%
           na.omit                         %>%
           droplevels()
     ## keep rows with valid 'monCount'
 row.names(myData) <- 1:nrow(myData)
     ## Reset index after removing NA rows from foodData
 myData

```

## Random Sampling
## Get a random sample from the population myData df
```{r}

 n <- 30
 s_row_nums <- sample(1:nrow(myData), n)
 s_myData   <- myData[s_row_nums,]
 s_myData
 
```

## Test #1: Use t-test on monCount of the sample s_myData
```{r}

  s_monCount <- s_myData$monCount
  t.test(s_monCount)
      ## sample mean (X_mean) & the confidence interval (CI)
  CI_lower <- t.test(s_monCount)$conf.int[1]
  CI_upper <- t.test(s_monCount)$conf.int[2]
  p_value  <- t.test(s_monCount)$p.value
  
  cat ("  This 30 sample has...")   %>%   print()
  
  table(s_myData$id)
      ## show # of avocado & ice-cream in this sample of 30

  mu_myData <- mean(myData$monCount)
      ## compute the mean of population
  
  if ( (mu_myData >= CI_lower) & (mu_myData <= CI_upper) ) {
     cat("\n  Accept the t-test -- population mean mu ",
         mu_myData," is \n  within the confidence interval: ",
         CI_lower,"-",CI_upper,".     ")   %>%
     print()
  } else if (p_value < 0.05) {
     cat("\n  Reject based on small p-value of "
         ,p_value," < 0.05.\n")            %>%
     cat("\n  Reject t-test since mean_X is outside the CI ",
         "interval &\n small p-value of ", p_value,
         " < 0.05.     ")                  %>%
     print()
  } else print("\n   Reject t-test -- mu outside CI range.     ")
  
```

## Compute the mean of sample for avocado & ice cream
```{r}

  s_myData_id  <-   s_myData  %>%  group_by(id)
  s_myData_id  %>%  summarize(mean(monCount))
  
```
 
## Let us do a t-test of the difference of means. The Null hyposthesis is that there is no difference (Xmean-Ymean=0)
```{r}

 avocado   <- s_myData                  %>%
              filter(id=="avocado")     %>%
              select(monCount)
 
 ice_cream <- s_myData                  %>%
              filter(id=="ice-cream")   %>%
              select(monCount)
 
 t.test(avocado,ice_cream)

 ## Check if difference in mean of zero is in the CI
 ## Check if p_value is less than 0.05
 
  CI_lower <- t.test(avocado,ice_cream)$conf.int[1]
  CI_upper <- t.test(avocado,ice_cream)$conf.int[2]
  p_value  <- t.test(avocado,ice_cream)$p.value
 
  cat("  Accept or reject the null hyposthesis?     ")   %>%
  print()
 
  if ( (0 >= CI_lower) & (0 <= CI_upper) ) {
     cat("\n  Accept the null hypothesis based difference ",
         "mean of zero is within \n   the confidence interval: ",
         CI_lower, " - ", CI_upper,".     ")       %>%
     print()
  } else if (p_value < 0.05) {
     cat("\n  Reject the null hypothesis since 0 is outside the ",
         "confidence \n  interval & small p-value of ", 
         p_value," < 0.05.     ")                  %>%
     print()
  } else print("\n  Reject the null hypothesis -- ",
               " zero is outside CI range.     ")

```

## Population: difference of means
```{r}

 myData_id  <-   myData  %>%  group_by(id)
 myData_id  %>%  summarize(mean(monCount))
 
```
